### 2020-10-30 smoothing.tex
clean up and adding testing strategies

Note that the equation for the untested susciptable compartment is $d S_u/dt = -\Lambda S_u - F_S S_u + \omega S_n$, where $\Lambda$ is FoI, $F_Z$ in general form is the weighted testing rate defined as $F_Z=\sigma W_Z$, and $\sigma$ is the testing rate. Let $W = W_S S_u + W_I I_u + W_R R_u$ be the weighted number of people available for tests.

\section{On testing rate}

The testing rate, $\sigma$, should be formulated such that people from the \_u compartments will not be tested if they're not there.

- The initial approache and the singularity issue;
$\sigma = \frac{\rho N_0}{W}$ was used initially. The issue was that the population in $S$ compartments appeared to blow up when the DFE is achieved. This is  once the only untested people are susceptibles, the FoI will become $\Lambda=0$, testing rate $F_s=\rho N_0/S_u$. Thus, eq(1) of the model will be $d S_u/dt = - \rho N_0 + \omega S_n$ which is no longer dependent on $S_u$ and a linear rate of leaving the $S_u$ compartment.

- The secondary approache;
It was suggested that

One natural way is to define $$\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}.$$
Where  $\tau$ ($1/ \mathrm{day}$) is a rate for the maximum rate of testing the whole untested population. In general, we want to test at a rate of $\rho$ across the whole population. This won't always be possible. So we impose a maximum rate of $\tau$ per testable person. It is consistent to think of both $\tau$ and $\rho$ as pure rates, but it might be clearer to think of $\rho$ as tests per capita per unit time, and $\tau$ as tests per testable person per unit time. It's not that we're switching through time, it's that we're imposing both of these as limitations. Note that, in general, since $\tau \gg \rho$, thus $\sigma=\rho \frac{\tau N0}{(\tau+\rho)N_0} \approx \rho \frac{\tau N0}{(\tau)N_0} = \rho$. This generally collapses to the original form. On the other hand, when $W$ is super-small, however, it collapses instead to $\tau$.
  In other words, at the beginning, we expect the answer to be close to $\rho N_0$ since $\tau$ should be very fast. Once $W$ becomes small, the limitation imposed by $\tau W$ will become important.

\section{ On the testing strategies}

- The testing weights, $W_S, W_I$ and $W_R$, are representing different testing strategies.
- The only thing that matters is the relative proportions of W's.
- Right now there are no restictions on the weights. If we impose $\sum_Z W_Z=1$, we are just reducing a degree of freedom. i.e., if we increase the $W_I$, the other two weights will decrease.

- Two things to explore are: 
(i) Random testing; this scenariocan be explored when $W_S=W_I=W_R$.

(ii) Testing and tracing; this scenariocan be explored when $W_S \ll W_I$ and for example $W_I=W_R=1$.