### 2020-12-08 codes/sir_macpan.Rnw
macpan/SIR equivalence

\documentclass{article}

\begin{document}
By Ali Gharouni,
<<date,echo=F>>=
Sys.time()
@
This is my a test for MacPan package.

\section{Install and Play}

<<install the package, echo=F  >>=
# install remotes package if necessary:
while (!require(remotes)) {
    install.packages("remotes")
}
## install development version of bbmle:
if (!require("bbmle") || packageVersion("bbmle") < "1.0.23.5") {
    remotes::install_github("bbolker/bbmle")
}
## install the target package and all its dependencies:
remotes::install_github("bbolker/McMasterPandemic",
                        dependencies = TRUE,
                        build_vignettes = TRUE
)
library(McMasterPandemic)
library(ggplot2); theme_set(theme_bw())
library(cowplot)

library(directlabels)

# params1 <- read_params("ICU1.csv")
# folder <- system.file("params", package="McMasterPandemic")
# list.files(folder)
# knitr::kable(describe_params(params1))

# knitr::kable(round(t(summary(params1)),2))
# knitr::kable(round(t(get_R0(params1, components=TRUE)),2))
@

<<6: MacPan Params , echo=F>>=
library(zoo)
library(tidyverse)

params <- read_params("ICU1.csv")

knitr::kable(round(t(summary(params)),2))
knitr::kable(round(t(get_R0(params, components=TRUE)),2)) 
@

<<Ali:SIR , echo=F>>=
# A simple SIR model
library(deSolve)
unpack <- McMasterPandemic::unpack

sir.mod <- function(time,state,params){
    unpack(as.list(c(state,params)))
    dS.dt <- -beta*S*I/N0
    dI.dt <-  beta*S*I/N0-gamma*I
    dR.dt <-  gamma*I
    # return the rate of change
    dxdt <- c(dS.dt,dI.dt,dR.dt)
    return(list(dxdt))
}
sir.params <- c(N0=10^6, beta=0.5,gamma=1/7)
class(sir.params) <- "params_pansim"
# Set the initial state
sir.state_init <- c(S=sir.params[["N0"]], 
                I=1,
                R=0)
##set the number of days for SIR simulation as simulation time of MacPan
startdate <- as.Date("01-03-2020")
enddate <- as.Date("01-09-2020")
d <- seq(0,as.numeric(enddate-startdate),by=1)

sir.R0 <- sir.params[["beta"]]/sir.params[["gamma"]]

sir.out <- as.data.frame(
  ode(
    func=sir.mod,
    y=sir.state_init,
    times= d, 
    parms=update(sir.params)
    ))

sir.out2 <- sir.out %>%
  pivot_longer(c(S,I,R), names_to = "compartment", values_to = "value")

gg1 <- (ggplot(data=sir.out2, aes(x=time,y=value,col=compartment))
    + geom_line()
    # + scale_y_log10(limits=c(0.1,sir.params[["N0"]]))
)
direct.label(gg1,"last.bumpup")

@

<<Try MacPan to recover the SIR result, echo=FALSE>>=
## update MacPan's params according to the SIR params
##
big_rate <- 20
params <- update(params, c(beta0=sir.params[["beta"]],
                           alpha=0, ## no asymptomatic
                           sigma=big_rate, ## very short latent period
                           ## would like this to -> Inf but can't
                           gamma_a=0,
                           mu=1,  ## all cases mild
                           gamma_m=sir.params[["gamma"]],
                           gamma_s=0,
                           ## try to skip presymptomatic period
                           gamma_p=big_rate
                           ))

pp <- params
## pp <- fix_pars(params, target = c(R0 = sir.R0)) #not sure if I need to update Gbar=6?
# Q? should I update beta0 again? 
state <- make_state(params=pp)
state[] <- 0
state[["S"]] <- 1e6
state[["Im"]] <- 1
# state <- update(state,c(S=sir.params[["N0"]],E=0,Ip=0,Im=0))

summary(pp)
sim0 <- run_sim(pp,state,
                start_date=startdate,
                end_date=enddate,
                use_ode=TRUE)

gg0 <- (ggplot(sim0,aes(x=date))
+ geom_point(aes(y=incidence))
)
print(gg0)

# combining the SIR and MacPan results also calculating incidences fr the SIR model
sim_combined <- sim0 %>% mutate(sir.S=sir.out[["S"]],
                                sir.I=sir.out[["I"]],
                                sir.S=sir.out[["R"]],
                                sir.incidence=sir.params[["beta"]]*sir.out[["S"]]*sir.out[["I"]])

gg01 <- (ggplot(sim_combined,aes(x=date))
+ geom_point(aes(y=I))
+ geom_lines(aes(y=sir.I), color='red')
)
print(gg01)
print(gg01 + scale_y_log10())
@

\end{document}