### 2021-05-14 note/cohort.tex
starting cohort-based framework

\documentclass[12pt]{article}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{grffile}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amssymb,amsmath,amsthm}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage[nameinlink,capitalize]{cleveref}
\usepackage{cleveref}
\usepackage[margin=1in]{geometry}
\usepackage{lineno}\renewcommand\thelinenumber{\color{gray}\arabic{linenumber}}
\usepackage{pdflscape}
\usepackage{enumerate}

\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}

\usepackage{xspace}

\newcommand{\fref}[1]{Fig.~\ref{#1}}
\newcommand{\Rlogo}{R\xspace}
\newcommand{\percap}{\emph{per capita}\xspace}
\newcommand{\Rnot}{\ensuremath{\mathcal{R}_0}}
\newcommand{\covid}{COVID-19\xspace}
\newcommand{\pro}[1][]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
\newcommand\pder[2][]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder[x]{y}

\newcommand{\comment}{\showcomment}
\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
\newcommand{\nocomment}[3]{}

\theoremstyle{definition} % amsthm only
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}

\bibliographystyle{apalike}

\title{The Paradigm of Cohort Equations for the SIR  Model With Testing and Isolation}

\begin{document}
\maketitle
\linenumbers

\section{Introduction}

\section{Ideas}
This is follwoing the work of \cite{champredon2018equivalence}.

In Watmough 2002's work matrix form of the I compartment $I'=(-V)I$, where I is n-by-1 and $V$ is n-by-1.

In the context of our SIR model with testing, $I^T=[I_u,I_n,I_p,I_c]$ where $T$ is the transposed operator.
And the solution is given by $I(\tau) = \exp(-V\tau) I(0)$, where $\tau$ is in the infection-time scale.
$K(\tau)=\Rnot g(\tau)= F I(\tau)$,
where $F$ is n-by-n matrix of new infections, $I(0)=[1,0,0,0]$.

$\exp(-V\tau)$ is the probability of being infected and stay infectious at time $\tau$.

$\exp(-V\tau)=\sum_{i=0}^\infty \tau^n (-V)^n/n!$.

\bibliography{../SIRlibrary}
\end{document}