### 2021-07-19 manuscript_BMB.tex

\RequirePackage{fix-cm}

\documentclass[smallextended]{svjour3}       % onecolumn (second format)

\smartqed  % flush right qed marks, e.g. at end of proof

\usepackage{graphicx}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{grffile}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amssymb,amsmath}
\usepackage{xcolor}
\usepackage{xspace}
% \usepackage[nameinlink,capitalize]{cleveref}
% \usepackage{cleveref}
\usepackage[margin=1in]{geometry}
\usepackage{lineno}\renewcommand\thelinenumber{\color{gray}\arabic{linenumber}}
\usepackage{pdflscape}
\usepackage{enumerate}
% \usepackage{authblk}
\usepackage{ragged2e}
\usepackage{array}
%
\hyphenation{co-ro-na-vi-rus}
\input{codes/modeldefs.tex}

% please place your own definitions here and don't use \def but
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\newcommand{\fref}[1]{Fig.~\ref{#1}}
\newcommand{\appref}[1]{Appendix~\ref{app:#1}}
% \newcommand{\Rlogo}{R\xspace}
\newcommand{\percap}{\emph{per capita}\xspace}
\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
\newcommand{\covid}{COVID-19\xspace}
\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
\newcommand{\pder}[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}
\newcommand{\testing}[1]{\ensuremath{{\cal T}_{#1}}\xspace}
\newcommand{\testinghat}[1]{\ensuremath{\hat{\cal T}_{#1}}\xspace}
\newcommand{\testtarget}{\ensuremath{w\_{IS}}}
\newcommand{\oldtesttarget}{\ensuremath{\frac{w_I}{w_S}}\xspace}
\newcommand*\subtxt[1]{_{\textnormal{#1}}}
%% Roman subscript
\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}

%% table stuff
\renewcommand{\arraystretch}{1.5}

% Insert the name of "your journal" with
\journalname{Bulletin of Mathematical Biology}

\begin{document}

\title{Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model
\thanks{Grants or other notes
about the article that should go on the front page should be
placed here. General acknowledgments should be placed at the end of the article.}
}
% \subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Ali Gharouni         \and
        F.M. Abdelmalek \and 
        David J. D. Earn \and
        Jonathan Dushoff \and
        Benjamin M. Bolker
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{Ali Gharouni \at
              Department of Mathematics and Statistics, McMaster University, Hamilton, Canada \\
              Tel.: +100-00-000000\\
              Fax: +100-00-000000\\
              \email{agharoun@uottawa.ca}\\
              \emph{Present address:}
              Department of Mathematics and Statistics, University of Ottawa, Ottawa, Ontario \\%  if needed\\
              \email{agharoun@uottawa.ca}\\
           \and
           F.M. Abdelmalek \at
              Department of Mathematics and Statistics, McMaster University, Hamilton, Canada \\
              \email{abdelf1@mcmaster.ca}\\  
           \and
           David J. D. Earn \at
              Department of Mathematics and Statistics, McMaster University, Hamilton, Canada\\
              Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Canada\\
              \email{earn@math.mcmaster.ca}\\
           \and
          Jonathan Dushoff \at
              Department of Biology, McMaster University, Hamilton, Canada\\
              Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Canada\\
              South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch, Stellenbosch,               South Africa\\
              \email{dushoff@mcmaster.ca} \\
           \and
          Benjamin M. Bolker \at
              Department of Mathematics and Statistics, McMaster University, Hamilton, Canada \\
              Department of Biology, McMaster University, Hamilton, Canada\\
              Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Canada\\
              \email{bolker@mcmaster.ca}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
\input{abstract.tex}
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
\subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
\input{introduction.tex}

\section{Methods}
\label{meth}
\input{methods.tex}

\section{Results}
\label{resu}
\input{results.tex}

\section{Discussion}
\label{disc}
\input{discussion.tex}

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

% Authors must disclose all relationships or interests that 
% could have direct or potential influence or impart bias on 
% the work: 
%
% \section*{Conflict of interest}
%
% The authors declare that they have no conflict of interest.

% BibTeX users please use one of
\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
\bibliography{SIRlibrary}   % name your BibTeX data base

\clearpage
\appendix
\section{Appendix}
\label{appe}
\input{appendix.tex}

\end{document}